---
output: html_document
---
## Data Generation Excercise

### Let's start by reading in the data. I'll call it `duncan`.

```{r}
duncan <- read.table("http://socserv.mcmaster.ca/jfox/Books/Companion/data/Duncan.txt")
```

## Now to answer some questions about the data:

### Summarize the data: 

* How many variables are there?  
I'm reasonably sure it's `r ncol(duncan)`.   

* How many cases?  
I'd like to say `r nrow(duncan)`, but I'm somewhat concerned that I may not have an adequate understanding of the term "case."

### Look at subsets of the data:

* What is the value of the education variable for the 35th case?  
  I suppose it's `r duncan[35,3]`.
* What is the mean of the education variable?  
  It should be `r mean(duncan$education)`, though for presentation it probably looks better as something like `r round(mean(duncan$education),1)`.
* How many different types of occupations are there?  
  I can't figure out a particularly pretty way to calculate this, but `r length(unique(duncan[,1]))` seems to be right.
* Which cases in the data have a prestige value greater than 85?  
  It took a while, but I found a solution I like that doesn't require any extra steps: `r rownames(duncan[which(duncan$prestige >85),])`.  
  It also just occured to me that the question might mean the number of the cases, in which case it's much easier: `r which(duncan$prestige > 85)`.
* View the value of all cases of the prestige variable that are greater than 85
  If you just want the list of values, that's easy: `r duncan$prestige[which(duncan$prestige > 85)]`.
* View the dataset for all cases in which prestige is higher than 85
  This is a little tougher, but I think this works:
  ```{r}
  duncan[which(duncan$prestige > 85),]
  ```
  Unfortunately, the output format is horrible outside the terminal.