---
output: pdf_document
---
## Canned Data Report
##### Edward Hernández
I'm going to complete the canned data report assignment from Fall 2014 (again) to brush up on my R skills.

Obviously, let's start by reading in some data.
```{r}
mturk.stress <- read.csv("mturkstress.csv",header=TRUE)
```

Now let's have some fun with it!

### Variables

#### How many subjects were in each of the conditions of the experiment?

The experiment had four treatment conditions: a control, a stress condition, a political stress condition, and a contentment condition. Participants were evenly distributed among these conditions.

```{r}
treatmenttable <- table(mturk.stress$treatmentgroup)
rownames(treatmenttable) <- c("Control","Stress","Politics","Content")
treatmenttable
```

#### What is the Mean of the Social Interaction Anxiety Scale variable?

```{r}
mean(na.omit(mturk.stress$saisscale))
```

#### Create a new variable called “partystrength” and make a table of its distribution.
##### It should take these values: `0` if the respondent is a true Independent, `1` if the respondent leans toward the Democrats or Republicans, `2` if the respondent is a weak Democrat or Republican, and `3` if the respondent is a strong Democrat or Republican. 


creating a null variable, running a loop to write in values based on party answers
```{r}
mturk.stress$partystrength <- NA
i <- 1926
while (i >= 1) {
	if(is.na(mturk.stress[i,23])==TRUE) {
		i <- i-1
	} else if(mturk.stress[i,23]==1) {
		if(is.na(mturk.stress[i,25])==TRUE) {
		} else if(mturk.stress[i,25]==1) {
			mturk.stress[i,137] <- 3
		} else if(mturk.stress[i,25]==2) {
			mturk.stress[i,137] <- 2
		}
		i <- i-1
	} else if(mturk.stress[i,23]==2) {
		if(is.na(mturk.stress[i,25])==TRUE) {
		} else if(mturk.stress[i,25]==1) {
			mturk.stress[i,137] <- 3
		} else if(mturk.stress[i,25]==2) {
			mturk.stress[i,137] <- 2
		}
		i <- i-1
	} else if(mturk.stress[i,23]==3) {
		if(is.na(mturk.stress[i,24])==TRUE) {
		} else if(mturk.stress[i,24]==1) {
			mturk.stress[i,137] <- 1
		} else if(mturk.stress[i,24]==2) {
			mturk.stress[i,137] <- 1
		} else if(mturk.stress[i,24]==3) {
			mturk.stress[i,137] <- 0
		}
		i <- i-1
	}
}
print(i)

partystrength.matrix <- matrix(table(mturk.stress$partystrength),ncol=1,nrow=4)
rownames(partystrength.matrix) <- c("True Independent","Leaning Independent","Weak Partisan", "Strong Partisan")
colnames(partystrength.matrix) <- c("n")

partystrength.matrix
```

#### Create an index variable for the “political engagement scale”.
##### Recode any of the six variables that are reverse coded and create an additive index.

reverse coding polengage2, polengage4, and polengage5 with a while loop#
```{r}
i <- 1926
while (i >= 1) {
      if(is.na(mturk.stress[i,28])==TRUE) {
	  }
	  else if(mturk.stress[i,28]==5){
	  mturk.stress[i,28] <- 1}
	  else if(mturk.stress[i,28]==4){
	  mturk.stress[i,28] <- 2}
	  else if(mturk.stress[i,28]==2){
	  mturk.stress[i,28] <- 4}
	  else if(mturk.stress[i,28]==1){
	  mturk.stress[i,28] <- 5}
	  
if(is.na(mturk.stress[i,30])==TRUE) {
	  }
	  else if(mturk.stress[i,30]==5){
	  mturk.stress[i,30] <- 1}
	  else if(mturk.stress[i,30]==4){
	  mturk.stress[i,30] <- 2}
	  else if(mturk.stress[i,30]==2){
	  mturk.stress[i,30] <- 4}
	  else if(mturk.stress[i,30]==1){
	  mturk.stress[i,30] <- 5}
	  
if(is.na(mturk.stress[i,31])==TRUE) {
	  }
	  else if(mturk.stress[i,31]==5){
	  mturk.stress[i,31] <- 1}
	  else if(mturk.stress[i,31]==4){
	  mturk.stress[i,31] <- 2}
	  else if(mturk.stress[i,31]==2){
	  mturk.stress[i,31] <- 4}
	  else if(mturk.stress[i,31]==1){
	  mturk.stress[i,31] <- 5}
i <- i-1
}

#checking iteration
print(i)



#summing a vector of the polengage variables, and saving it as polengage
mturk.stress$polengage <- NA
i <- 1926
while (i >=1) {
if(is.na(mturk.stress[i,27])==TRUE){
}
else if(is.na(mturk.stress[i,28])==TRUE){
}
else if(is.na(mturk.stress[i,29])==TRUE){
}
else if(is.na(mturk.stress[i,30])==TRUE){
}
else if(is.na(mturk.stress[i,31])==TRUE){
}
else{
mturk.stress[i,138] <- sum(mturk.stress[i,27:32])
}
i <- i-1
}

#checking iteration completed
print(i)
```

#### Create seperate tables that show the distribution of three variables in the dataset.

creating and printing a table for born in the US
```{r}
borninUS.table <- table(mturk.stress$borninUS)
rownames(borninUS.table) <- c("Yes", "No")
borninUS.table

marital.table <- table(mturk.stress$marital)
rownames(marital.table) <- c("Single","Married","Widowed","Divorced","Other")
marital.table

occupation.table <- table(mturk.stress$occupation)
rownames(occupation.table) <- c("Clerical","Professional","Self-Employed","Corporation","Other White Collar","Service","Custodial/Factory","Construction","Other Blue Collar","Homemaker","Student","Unemployed")
occupation.table
```

#### Identify two variables that may be related.

t.test would be good, but the hypotheses are in the doc

### Sample

#### What proportion of respondents finished the survey?

```{r}
length(mturk.stress$finished[mturk.stress$finished>0]) / nrow(mturk.stress)
```

#### What proportion of the sample is married?

```{r}
length(na.omit(mturk.stress$marital[mturk.stress$marital %in% c(2)])) / nrow(mturk.stress)
```

#### What proportion of the women in the sample is married?

```{r}
length(na.omit(mturk.stress$marital[mturk.stress$marital %in% c(2) & mturk.stress$male %in% c(2)])) / length(na.omit(mturk.stress$male[mturk.stress$male %in% c(2)]))
```

#### Are the proportions of men and women in the sample significantly different from each other?

```{r}
prop.test(table(mturk.stress$male))
```

#### Compare this dataset to the American National Election Study.
##### What proportion of people in 2008 were strong Democrats, weak Democrats, or leaned Democrat. Is that proportion significantly different from the equivalent proportion in our dataset? How does this affect our findings?


establishing
```{r}
mturk.stressdemstrong <- length(mturk.stress$polengage[mturk.stress$part1 %in% c(1) & mturk.stress$part3 %in% c(1)]) / nrow(mturk.stress)
mturk.stressdemweak <- length(mturk.stress$polengage[mturk.stress$part1 %in% c(1) & mturk.stress$part3 %in% c(2)]) / nrow(mturk.stress)
mturk.stressdemlean <- length(mturk.stress$polengage[mturk.stress$part1 %in% c(3) & mturk.stress$part2 %in% c(1)]) / nrow(mturk.stress)


mturk.stressdem.matrix <- matrix(c(mturk.stressdemstrong,mturk.stressdemweak,mturk.stressdemlean),ncol=1,nrow=3)
anesdem.matrix <- matrix(c(0.19,0.15,0.17),ncol=1,nrow=3)

demtestmatrix <- cbind(mturk.stressdem.matrix,anesdem.matrix)
demtestmatrix

prop.test(demtestmatrix)
```

### Hypothesis Testing

#### Did the political stress treatment cause respondents to report a lower likelihood of voting compared to respondents in the control group? First, be clear on what test you are running to assess this hypothesis. I will check your code to see if you implemented the test correctly. In your report, you should clearly explain your finding.

running a t-test to rule out the null hypothesis
t.test(mturk.stress$votelikely~mturk.stress$pol)

### Data Visualization

#### Make a bar chart of the proportion of supporters for each candidate.

creating proportion variables
```{r}
barack <- length(mturk.stress$votechoice[mturk.stress$votechoice %in% c(1)])/ nrow(mturk.stress) * 100
mitt <- length(mturk.stress$votechoice[mturk.stress$votechoice %in% c(2)]) / nrow(mturk.stress) * 100
other <- length(mturk.stress$votechoice[mturk.stress$votechoice %in% c(3)]) / nrow(mturk.stress) * 100


#creating a matrix of the variables

votechoice <- matrix(c(barack, mitt, other),nrow=1,ncol=3)


#plotting the matrix
barplot(votechoice,names.arg=c("Obama","Romney","Other"),beside=TRUE,col=c("dodgerblue4", "firebrick1", "darkgoldenrod4"), ylim=c(0,75),xlab="Candidates",font.lab="2",ylab="Proportion of Respondents",main="Vote Choice Proportions in the MTurk Dataset")
box()
```

#### Make a histogram of the age variable.

```{r}
#creating a histogram of age
hist(mturk.stress$age,xlab="Age",font.lab="2",main="Histogram of MTurk Participant Age",col=c("chartreuse4"))
box()
```

#### Make a correlation plot between the age variable and the response length variable, for all those people whose response is greater than 0. What is the correlation between those two variables (you will need to run a correlation test to get the answer to this)?

```{r}
#creating plot of age v response time and regression line
plot(na.omit(mturk.stress$age[mturk.stress$responselength>0]),na.omit(mturk.stress$responselength[mturk.stress$responselength>0]),xlab="Age",ylab="Response Length",yaxt="n",ylim=c(0,400),main="Correlation Between Respondent Age and Response Length")
abline(lm(na.omit(mturk.stress$responselength[mturk.stress$responselength>0])~na.omit(mturk.stress$age[mturk.stress$responselength>0])),col="red")

#running correlation test
cor.test(na.omit(mturk.stress$age[mturk.stress$responselength>0]),na.omit(mturk.stress$responselength[mturk.stress$responselength>0]))
```